\documentclass[12pt]{extarticle}

\input{../preamble.tex}

\fancyhead[R]{\textbf{Homework \#1}}
\setlength\parindent{0pt}

\begin{document}

\section*{Problem 1}

Express each of the following statements in the “If \ldots . , then \ldots . ” form. There are many possible correct answers. 
\begin{enumerate}[label=(\alph*)]
	\item You must eat your dinner if you want to grow. 
	\item Being a multiple of 12 is a sufficient condition for a number to be even. 
	\item It is necessary for you to pass your exams in order for you to obtain a degree. 
	\item A triangle is equilateral only if all its sides have the same length.
\end{enumerate}

\subsection*{Solution}

\begin{enumerate}[label=(\alph*)]
	\item \textbf{If} you want to grow, \textbf{then} you must eat your dinner
	\item \textbf{If} a number is a multiple of 12, \textbf{then} it is even
	\item \textbf{If} you want to obtain a degree, \textbf{then} you need to pass your exams
	\item \textbf{If} a triangle is equalateral, \textbf{then} all its sides are of equal length
\end{enumerate}

\section*{Problem 2}

Suppose that “Girls smell of roses” and “Boys have dirty hands” are true statements and that “The Teacher is always right” is a false statment. Which of the following are true? \textit{Hint: Label each of the given statements, and think about each of the following using connectives.} 

\begin{enumerate}[label=(\alph*)]
	\item If girls smell of roses, then the Teacher is always right. 
	\item If the Teacher is always right, then boys have dirty hands. 
	\item If the Teacher is always right or girls smell of roses, then boys have dirty hands.
	\item If boys have dirty hands and girls smell of roses, then the Teacher is always right.
\end{enumerate}

\subsection*{Solution}

Let $P$ = "Girls smell of roses", $Q$ = "Boys have dirty hands", and $R$ = "The Teacher is Always Right".

\begin{enumerate}[label=(\alph*)]
	\item $P \implies R$
	\begin{itemize}
		\item True $\implies$ False is \textbf{False}
	\end{itemize}
	\item $R \implies Q$
	\begin{itemize}
		\item False $\implies$ True is \textbf{True}
	\end{itemize}
	\item $(R\lor Q) \implies P$
	\begin{itemize}
		\item True $\implies$ True is \textbf{True}
	\end{itemize}
	\item $(Q\land P) \implies R$
	\begin{itemize}
		\item True $\implies$ False is \textbf{False}
	\end{itemize}
\end{enumerate}

\section*{Problem 3}

Write the negation (in words) of the following claim: \textit{If Jack and Jill climb up the hill, then they fall down and like pails of water.}

\subsection*{Solution}

Jack and Jill climb up the hill and dont fall down or dont like pails of water.

\section*{Problem 5}

\begin{enumerate}[label=(\alph*)]
	\item Rewrite the following sentence using the word ‘necessary.’ \\
		\textit{If I am to get a new bicycle, I must do my homework.}
	\item Rewrite the following sentence using the word ‘sufficient.’ \\
		\textit{The United States must play more soccer if it is to win the World Cup}
\end{enumerate}

\subsection*{Solution}

\begin{enumerate}[label=(\alph*)]
	\item Doing my homework is \textbf{neccessary} for me to get a new bicycle
	\item To win the world cup it is \textbf{sufficient} for the United States to play more soccer
\end{enumerate}

\section*{Problem 6}

\begin{enumerate}[label=(\alph*)]
	\item What are the converse and the contrapositive of the statements in the previous question? Write your answers in sentences, like the originals.
	\item What are the negations of the statements in the previous question? 
\end{enumerate}

\subsection*{Solution}

\subsubsection*{Part A}

\begin{itemize}
	\item Converse Statements
	\begin{enumerate}[label=(\alph*)]
		\item If I do my homework, then I can get a new bicycle
	\end{enumerate}
	\item Contrapositive Statements
	\begin{enumerate}[label=(\alph*)]
		\item If I do not do my homework, then I will not get a new bicycle 
	\end{enumerate}
\end{itemize}

\subsubsection*{Part B}
\begin{enumerate}[label=(\alph*)]
	\item I will get a new bicycle and not do my homework
\end{enumerate}

\section*{Problem 7}

Construct the truth tables for the propositions $P \lor (Q \land R)$ and $(P \lor Q) \land R$. Are they the same?

\subsection*{Solution}

\begin{minipage}[c]{0.5\textwidth}
	\centering
	\begin{tabular}{ccc | c || c}
		$P$ & $Q$ & $R$ & $Q\land R$ & $P \lor (Q \land R)$ \\\hline
		T & T & T & T & T \\ 
		T & T & F & F & T \\ 
		T & F & T & F & T \\ 
		T & F & F & F & T \\ 
		F & T & T & T & T \\ 
		F & T & F & F & F \\ 
		F & F & T & F & F \\ 
		F & F & F & F & F \\ 
	\end{tabular}
\end{minipage}
\begin{minipage}[c]{0.5\textwidth}
	\centering
	\begin{tabular}{ccc | c || c}
		$P$ & $Q$ & $R$ & $P \lor Q$ & $(P \lor Q) \land R$ \\\hline
		T & T & T & T & T \\ 
		T & T & F & T & F \\ 
		T & F & T & T & T \\ 
		T & F & F & T & F \\ 
		F & T & T & T & T \\ 
		F & T & F & T & F \\ 
		F & F & T & F & F \\ 
		F & F & F & F & F \\ 
	\end{tabular}
\end{minipage}
\vspace{1cm}

The propositions are \textbf{not the same}

\section*{Problem 10}

\begin{enumerate}[label=(\alph*)]
  \item Prove that $((P \lor Q) \land \lnot P) \land \lnot Q$ is a contradiction.
  \item Prove that $(\lnot P \land Q) \lor (P \land \lnot Q) \Longleftrightarrow \lnot(P \Longleftrightarrow Q)$ is a tautology.
\end{enumerate}

\subsection*{Solution}
\subsubsection*{Part A}

A proposition is a contradiction if for every input it is false.

\begin{table}[h!]
	\centering
	\begin{tabular}{cc || c|c|c|c||c}
		$P$ & $Q$ & $\lnot P$ & $\lnot Q$ & $P \lor Q$ & $(P \lor Q) \land \lnot P$ & $((P \lor Q) \land \lnot P) \land \lnot Q$ \\\hline
    T & T & F & F & T & F & F \\
    T & F & F & T & T & F & F \\
    F & T & T & F & T & T & F \\
    F & F & T & T & F & F & F \\
	\end{tabular}
\end{table}

Since for all posible inputs $P$ and $Q$ the proposition is false, it is a contradiction.

\newpage
\subsubsection*{Part B}

Let $H$ be the proposition in question; $H \Longleftrightarrow (\lnot P \land Q) \lor (P \land \lnot Q) \Longleftrightarrow \lnot(P \Longleftrightarrow Q) $

\begin{table}[h!]
	\centering
	% \footnotesize
	\begin{tabular}{cc || c|c|c|c||c}
		$P$ & $Q$ & $\lnot P \land Q$ & $P \land \lnot Q$ & $(\lnot P \land Q)\lor(P \land \lnot Q)$ & $\lnot(P \Longleftrightarrow Q)$ & $H$\\\hline
		T	& T & F & F & F & F & T \\ 
		T	& F & F & T & T & T & T \\ 
		F	& T & T & F & T & T & T \\ 
		F	& F & F & F & F & F & T
	\end{tabular}
\end{table}

Since $H$ is always true and is the original proposition, then $(\lnot P \land Q) \lor (P \land \lnot Q) \Longleftrightarrow \lnot(P \Longleftrightarrow Q)$ is a tautology.

% First note that the negation of $P \Longleftrightarrow Q$ is equivalent to $P\land \lnot Q$. Secondly, by the commutative law of logic
% \begin{align*}
% 	(\lnot P \land Q) \lor (P \land \lnot Q) &\Longleftrightarrow (\lnot P \land (P\land \lnot Q)) \lor (Q \land (P\land \lnot Q)) \\
% 	\intertext{By the associative and commutative law of logic, the second half can be further rewritten as}
% 	(\lnot P \land (P\land \lnot Q)) \lor (Q \land (P\land \lnot Q)) &\Longleftrightarrow((\lnot P \land P) \land \lnot Q) \lor ((Q \land \lnot Q)\land P)
% \end{align*}

\section*{Problem 14}

Suppose that the following statements are true: 

\begin{itemize}
  \item Every Pig likes mud. 
  \item If a creature cannot fly then it is not an astronaut. 
  \item A creature is an astronaut if it likes mud. 
\end{itemize}

Is it true that ‘Pigs can fly’? Explain your answer. 

\subsection*{Solution}

First we assume that pigs are considered creatures. Each statement can be rewritten as:

\begin{itemize}
	\item Every pig likes mud
	\item Cannot fly $\implies$ Is not an astronaut
	\item Likes mud $\implies$ Is an astronaut
\end{itemize}

By statement 1, every pig likes mud. By statement 3, any creature that likes mud is an astronaut. Therefore every pig is an astronaut. The contrapositive of statement 2 results in "If a creature is an astronaut, then it can fly". Since the contrapositive of a proposition is logically equivalent to the original proposition, this proposition is also true. Therefore, since every pig is an astronaut and if a creature is an astronaut, it can fly, it is true that all pigs can fly.

\end{document}
